Supporting Information: Unsupervised embedding of trajectories captures the latent structure of mobility

### 2022-12-05 paper/current/supp.tex
revision_second_round

With this simulated geography in place, we calculate the pairwise distance matrix $r_{ij}^{synthetic}$, which is then used to calculate a transition matrix between points, defined as $T_{ij} = M_i M_j f(r_{ij})$, with transition probabilities following the gravity law. The purpose of this transition matrix is to define the likelihood of movement between points (or \textit{locations}) in our simulated geography based on their mass and pairwise distances. We conduct this experiment with two different transition matrices, one based on cosine distance where $f(r_{ij})=e^{-\beta r_{ij}}$ and another with euclidean distance where $f(r_{ij})=r_{ij}^{-\alpha}$. For simplicity, we use $\beta=2$ and $\alpha=2$, though the analysis was repeated for multiple parameters in order to assess robustness.

@@

@@

>>>>

With this simulated geography in place, we calculate the pairwise distance matrix $r_{ij}^{synthetic}$, which is then used to calculate a transition matrix between points, defined as $T_{ij} = M_i M_j f(r_{ij})$, with transition probabilities following the gravity law. The purpose of this transition matrix is to define the likelihood of movement between points (or \textit{locations}) in our simulated geography based on their mass and pairwise distances. We conduct this experiment with two different transition matrices, one based on cosine distance where $f(r_{ij})=e^{-\beta r_{ij}}$ and another with euclidean distance where $f(r_{ij})=r_{ij}^{-\alpha}$. For simplicity, we use $\beta=2$ and $\alpha=2$.

@@

In order to evaluate the robustness of the outcome across hyperparameter settings, we repeat our synthetic benchmark (using cosine distance) varying values for the mass exponent, $k$ and distance exponent $\beta$. 
In Fig. \ref{fig:supp:synthetic_robust}) presents the average Pearson correlation between the synthetic and embedding distance over ten iterations with for different values of $k$ and $\beta$. 
This shows our findings to be robust. 
Neural embedding consistently recovers the distance from the synthetic trajectories across a broad range of hyperparameter values. 
A low correlation occurs only when the mass distribution is too heterogeneous ($k < 1.9$).
In a heterogeneous mass distribution, most of the trajectories will concentrate on just a few high-mass locations, regardless of distance, which makes learning a distance challenging.  Meanwhile, the exponent of the city population usually ranges between 2 and 3~\autocite{clauset2009power}.

@@

\newpage
%
% Figure - reverse CDF by scale
%
\begin{figure}[p!]
	\centering
	\includegraphics[width=0.8\textwidth]{\figdir/rebuttal/robustness_check.pdf}
	    \caption{\textbf{Neural embedding consistently recovers the underlying distance from trajectories across diverse hyperparameter settings.} We repeat our synthetic benchmark varying mass exponent $k$ and distance exponent $\beta$. The color of each cell represents the average Pearson correlation of the synthetic distance and embedding distances over ten repetitions with given $k$ and $\beta$.}
	\label{fig:supp:synthetic_robust}
\end{figure}
